---
title: "Rueda-Alaña Chick Early Data"
author: "Rodrigo Senovilla Ganzo"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)
library(clustree)
library(pvclust)
library(Hmisc)
library(corrplot)
library(ggcorrplot)
library(stringr)
library(biomaRt)
```

## 1to1 otrhologs

Obtain orthologs names, the substitution step will be done at the end though

## Mouse genes 1to1

```{r cars}
if(exists("ensembl")==F){
ensembl<-useMart("ensembl")}
datasets <- listDatasets(ensembl)
if(exists("mouse")==F){
mouse<-useDataset("mmusculus_gene_ensembl", mart=ensembl)}
attributes = c("external_gene_name", "ensembl_gene_id", "ggallus_homolog_ensembl_gene","ggallus_homolog_perc_id_r1", "ggallus_homolog_orthology_type", "ggallus_homolog_subtype", "ggallus_homolog_perc_id")
mouse_genes <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"), mart = mouse)
for(i in 1:length(mouse_genes[,1])){
  if(mouse_genes[i,1]==""){
    mouse_genes[i,1]<-mouse_genes[i,2]
  }
}
orth.mouse = getBM(attributes, filters="with_ggallus_homolog",values=TRUE, mart = mouse, bmHeader=FALSE)
orth.mouse<-orth.mouse[which(orth.mouse$ggallus_homolog_orthology_type=="ortholog_one2one"),]
ensembl.mouse<- orth.mouse[,2]
external.mouse<-orth.mouse[,1]
rownames(orth.mouse)<-ensembl.mouse
TFdata_mouse <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"),filters = 'go', values = c('GO:0006366', 'GO:0000981', 'GO:0003700', 'GO:0006383', 'GO:0000995', 'GO:0001228', 'GO:0001227'), mart = mouse)
```

## Chicken genes 1to1

```{r cars}
if(exists("ensembl")==F){
ensembl<-useMart("ensembl")}
datasets <- listDatasets(ensembl)
if(exists("chicken")==F){
chicken<-useDataset("ggallus_gene_ensembl", mart=ensembl)}
attributes = c("external_gene_name", "ensembl_gene_id", "mmusculus_homolog_ensembl_gene","mmusculus_homolog_perc_id_r1", "mmusculus_homolog_orthology_type", "mmusculus_homolog_subtype", "mmusculus_homolog_perc_id")
chicken_genes <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"), mart = chicken)
for(i in 1:length(chicken_genes[,1])){
  if(chicken_genes[i,1]==""){
    chicken_genes[i,1]<-chicken_genes[i,2]
  }
}
orth.chicken = getBM(attributes, filters="with_mmusculus_homolog",values=TRUE, mart = chicken, bmHeader=FALSE)
orth.chicken<-orth.chicken[which(orth.chicken$mmusculus_homolog_orthology_type=="ortholog_one2one"),]
ensembl.chicken<- orth.chicken[,2]
external.chicken<-orth.chicken[,1]
rownames(orth.chicken)<-ensembl.chicken
TFdata_chicken <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"),filters = 'go', values = c('GO:0006366', 'GO:0000981', 'GO:0003700', 'GO:0006383', 'GO:0000995', 'GO:0001228', 'GO:0001227'), mart = chicken)
```


#Indentities

This step will be removed in the future, as the identities will be well stablished beforehand and we will know what each cluster stands for. 

```{r , echo=FALSE}
me9 <- readRDS("../mouse_e9/me9_sct5f.rds")
era <-readRDS("../chicken_e3/ce3_4f.rds")
```

```{r , echo=FALSE}
me9[['seurat_clusters']]<-me9[['me9_sct5_res2.8']]
era[['seurat_clusters']]<-era[['ce3_4']]
Idents(me9) <- "seurat_clusters"
Idents(era) <- "seurat_clusters"
```

```{r cars, warning=FALSE}
era_orth<-era
era.data<-GetAssayData(object = era, assay = "RNA")
era.data_ensembl<-merge(x=chicken_genes[,1:2], y=era.data, by.x= "external_gene_name", by.y= "row.names")
era.data_orth<-merge(x=orth.chicken[,2:3], y=era.data_ensembl, by.x= "ensembl_gene_id", by.y= "ensembl_gene_id")
era.data_orth<-era.data_orth[,-c(1,3)]
era.data_orth<-merge(x=mouse_genes, y=era.data_orth, by.x = "ensembl_gene_id", by.y = "mmusculus_homolog_ensembl_gene")
era.data_orth$external_gene_name<-sub('[.]','-',make.names(era.data_orth$external_gene_name, unique=T))
rownames(era.data_orth)<-toupper(era.data_orth$external_gene_name)
era.data_orth<-era.data_orth[,-c(1,2)]
era.data_orth<-Matrix(as.matrix(era.data_orth),sparse=TRUE)
mouse_gene_assay <- CreateAssayObject(era.data_orth)
era_orth[['RNA']] <- mouse_gene_assay
```


##Integration of datasets

```{r , echo=FALSE}
DefaultAssay(era_orth)<-"RNA"
chick.list <- SplitObject(era_orth, split.by="orig.ident")
chick.list <- lapply(X = chick.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("S.Score", "G2M.Score","percent.mt", "z.Score", "w.Score", "percent.hist"))
features <- SelectIntegrationFeatures(object.list = chick.list)
chick.list <- PrepSCTIntegration(object.list = chick.list, anchor.features = features)
chick.list <- lapply(X = chick.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = chick.list, reference = c(1, 2), reduction = "cca", dims = 1:50, normalization.method = "SCT", anchor.features = features)
era_orth <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```


```{r cars, warning=FALSE}
me9_orth<-me9
me9.data<-GetAssayData(object = me9, assay = "RNA")
orth.mouse[,1]<-toupper(orth.mouse[,1])
me9.data_orth<-merge(x=orth.mouse[,1:2], y=me9.data, by.x= "external_gene_name", by.y= "row.names")
rownames(me9.data_orth)<-me9.data_orth$external_gene_name
me9.data_orth<-me9.data_orth[,-c(1,2)]
me9.data_orth<-Matrix(as.matrix(me9.data_orth),sparse=TRUE)
mouse_gene_assay <- CreateAssayObject(me9.data_orth)
me9_orth[['RNA']] <- mouse_gene_assay
```

```{r}
DefaultAssay(me9_orth)<-"RNA"
me9_orth.list <- SplitObject(me9_orth, split.by = "orig.ident")
me9_orth.list <- lapply(X = me9_orth.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("percent.mt", "G2M.Score", "S.Score"))
features <- SelectIntegrationFeatures(object.list = me9_orth.list)
me9_orth.list <- PrepSCTIntegration(object.list = me9_orth.list, anchor.features = features)
me9_orth.list <- lapply(X = me9_orth.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = me9_orth.list, reduction = "cca", dims = 1:50, normalization.method = "SCT", anchor.features = features)
me9_orth <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```
#Markers

Although this would have alredy been done in the past, it is easier to get it again than save all gene markers. Or maybe not, it could be removed then. 

```{r, echo=F}
era.markers<-FindAllMarkers(era_orth,only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
me9.markers<-FindAllMarkers(me9_orth, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```
#Top gene markers

Select the most differentially expressed genes, the top 400 of each cluster ordered by |avg_log2FC| and filtered by p_val_adj < 10^(-9). This was set based on Colquitt, but it could be changed. 

```{r, echo=F}
era.markers_top <-era.markers %>% group_by(cluster) %>% filter(p_val_adj < 10^(-9)) %>% arrange(abs(avg_log2FC)) %>% top_n(n=400) %>% dplyr::select(gene) 
table(era.markers_top$cluster)
me9.markers_top<-me9.markers %>% group_by(cluster) %>% filter(p_val_adj < 10^(-9)) %>% arrange(abs(avg_log2FC)) %>% top_n(n=400) %>% dplyr::select(gene) 
table(me9.markers_top$cluster)
```


#Top Markers Subset - ERA

```{r, echo=F}
era_top<-era_orth
era_top
era_top[["SCT"]]@counts<-era_top[["SCT"]]@counts[era.markers_top$gene,]
era_top[["SCT"]]@data<-era_top[["SCT"]]@data[era.markers_top$gene,]
era_top[["SCT"]]@meta.features[era.markers_top$gene,'cluster']<-era.markers_top$cluster
era_top
```
#g coefficient - ERA

```{r, echo=F}
Idents(era_top)<-"seurat_clusters"
era_cluster.averages <- AverageExpression(era_top, assays = "SCT")
Idents(era_top)<-"SCT_snn_res.0"
era_averages <- AverageExpression(era_top, assays = "SCT")
era_g<-era_cluster.averages$SCT/era_averages$SCT[,1]
head(era_g)
```

#Top Markers Subset - me9


```{r, echo=F}
me9_top<-me9_orth
me9_top
me9_top[["SCT"]]@counts<-me9[["SCT"]]@counts[me9.markers_top$gene,]
me9_top[["SCT"]]@data<-me9[["SCT"]]@data[me9.markers_top$gene,]
me9_top[["SCT"]]@meta.features[me9.markers_top$gene,'cluster']<-me9.markers_top$cluster
me9_top
```
#g coefficient - me9

```{r, echo=F}
Idents(me9_top)<-"seurat_clusters"
me9_cluster.averages <- AverageExpression(me9_top, assays = "SCT")
Idents(me9_top)<-"SCT_snn_res.0"
me9_averages <- AverageExpression(me9_top, assays = "SCT")
me9_g<-me9_cluster.averages$SCT/me9_averages$SCT[,1]
head(me9_g)
```


#Correlation matrix

Although the each species matrix only includes ortholog, we keep for comparisons only those genes presents in both datasets. 

#Chicken vs Mice

```{r, echo=F}
era_ct<- c(
"MB/DIEN-RP
HB-RP
HEM
AMP
OV
aTEL
bTEL
tHT
pHT
P3
aP2
bP2
P1
PG
aIT
bMB
pabMB
a'abMB
aMB
pIT
RL
aR1-2
aR3-8
abR0-8
bR0-8
DIEN-FP
MB-FP
IT-FP
HB-aFP
HB-bFP
TAL1 N
MOT N
TAL2 N
LHX9 N
LHX3 N
SPAL N
EI N
LI N"
)
era_ct<-str_split(era_ct, '\n')[[1]]
era_ct<-unlist(era_ct)
```

```{r, echo=F}
me9_ct<- c(
"MB/DIEN-RP
HB-RP
HEM
AMP
OV
TEL1
TEL2
SPAL
tHT
ATO
pHT
P3
P1-2
aIT
aMB
bMB
paIT
RL
EGL
aR1
aR2
aR3
aR4
aR5
abR1-2
abR3
abR4
bR0-2
bR3-5
MB-FP
IT-FP
HB-FP
MOT N
TAL2 N
MSX3 N
LHX N
LI N")
me9_ct<-str_split(me9_ct, '\n')[[1]]
me9_ct<-unlist(me9_ct)
```

```{r, echo=F}
era_gf<-as.matrix(era_g[rownames(era_g) %in% rownames(me9_g),])
era_gf<-era_gf[,era_ct]
colnames(era_gf)<-lapply(colnames(era_gf), function(x) paste("C", x, sep="-"))
me9_gf<-as.matrix(me9_g[rownames(me9_g) %in% rownames(era_g),])
me9_gf<-me9_gf[,me9_ct]
colnames(me9_gf)<-lapply(colnames(me9_gf), function(x) paste("M", x, sep="-"))
table(rownames(me9_gf) %in% rownames(era_gf))
#Step7: Merge Expression Tables
 geTable = merge(era_gf,me9_gf, by='row.names', all=F)
  rownames(geTable) = geTable$Row.names
  geTable = geTable[,2:ncol(geTable)]
pb_mc.cor<-cor(geTable, method = "spearman")
pb_mc.cor
```

```{r, echo=F}
#8b:  Shuffle data
  shuffled.cor.list = list()
  pb_mc   <- txtProgressBar(1, 100, style=3)
nPermutations <- 100
  for (i in 1:nPermutations){
    shuffled = apply(geTable[,1:ncol(era_gf)],1,sample)
    shuffled2 = apply(geTable[,(ncol(era_gf)+1):ncol(geTable)],1,sample)
    shuffled = cbind(t(shuffled),t(shuffled2))
    shuffled.cor = cor(shuffled,method="spearman")
    shuffled.cor.list[[i]] = shuffled.cor
    rm(list=c('shuffled','shuffled2','shuffled.cor'))
    if ((i %% 100) ==0){
      setTxtProgressBar(pb_mc, (i*100)/nPermutations)
    }
  }
  
  p.value.table = matrix(ncol=ncol(geTable), nrow = ncol(geTable))
  rownames(p.value.table) = colnames(geTable)
  colnames(p.value.table) = colnames(geTable)
  
  shuffled.mean.table = matrix(ncol=ncol(geTable), nrow = ncol(geTable))
  rownames(shuffled.mean.table) = colnames(geTable)
  colnames(shuffled.mean.table) = colnames(geTable)
  
  a = combn(1:ncol(geTable),2)
  for (i in 1:ncol(a)){
    cor.scores = sapply(shuffled.cor.list,"[",a[1,i],a[2,i])
    shuffled.mean.table[a[1,i],a[2,i]] = mean(cor.scores)
    shuffled.mean.table[a[2,i],a[1,i]] = mean(cor.scores)
    p.value = mean(abs(cor.scores)>=abs(pb_mc.cor[a[1,i],a[2,i]]))
    p.value.table[a[1,i],a[2,i]] = p.value
    p.value.table[a[2,i],a[1,i]] = p.value
    rm(list=c('cor.scores','p.value'))
    setTxtProgressBar(pb_mc, (i*100)/ncol(a))
  }
p.value.table[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]
  neg.log10.p = -log10(p.value.table)
```
```{r}
  #step9 "Overlap in Markers"
  #for all pairs of cell-types generate list of genes that are at least 1.5x avg in both cells
  
  #from above a = combn(1:ncol(geTable),2)
  marker.overlap.list = list()
  for (i in 1:ncol(a)){
    datasubset = cbind(geTable[,a[1,i]],geTable[,a[2,i]])
    markers = rownames(geTable[datasubset[,1]>1.5 & datasubset[,2]>1.5,])
    marker.overlap.list[[i]] = markers
    names(marker.overlap.list)[i] = paste(colnames(geTable)[a[1,i]], colnames(geTable)[a[2,i]],sep='_')
    rm(list=c('datasubset','markers'))
  }
```

```{r}
colnames(pb_mc.cor)<-str_sub(colnames(pb_mc.cor), start = 3)
rownames(pb_mc.cor)<-str_sub(rownames(pb_mc.cor), start = 3)
```


```{r, fig.height=7, fig.width=10, echo=F}
#plot it 
corrplot::corrplot(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], order="original", tl.pos="lt", method="color", tl.col="black",cl.lim=c(min(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]),max(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)])), is.corr=F,tl.cex=0.7, sig.level=(0.05),insig="pch", pch=5, p.mat=p.value.table[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], pch.cex=0.25,pch.col="black",mar=c(3,1,5,1),cl.align.text="l")
corrplot::corrplot(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], order="original", tl.pos="lt", method="color", tl.col="black",cl.lim=c(min(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]),max(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)])), is.corr=F,tl.cex=0.7, sig.level=(0.05),insig="pch", pch=5, p.mat=p.value.table[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], pch.cex=0.25,pch.col="black",mar=c(3,1,5,1),cl.align.text="l", addgrid.col = "black")
```
```{r, fig.height=7, fig.width=10, echo=F}
#plot it 
pdf("corr_me9_vs_ce3_subsetprev.pdf")
corrplot::corrplot(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], order="original", tl.pos="lt", method="color", tl.col="black",cl.lim=c(min(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]),max(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)])), is.corr=F,tl.cex=0.7, sig.level=(0.05),insig="pch", pch=5, p.mat=p.value.table[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], pch.cex=0.25,pch.col="black",mar=c(3,1,5,1),cl.align.text="l")
corrplot::corrplot(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], order="original", tl.pos="lt", method="color", tl.col="black",cl.lim=c(min(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]),max(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)])), is.corr=F,tl.cex=0.7, sig.level=(0.05),insig="pch", pch=5, p.mat=p.value.table[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], pch.cex=0.25,pch.col="black",mar=c(3,1,5,1),cl.align.text="l", addgrid.col = "black")
dev.off()
```
```{r, fig.height=7, fig.width=10, echo=F}
#plot it 
p_inv<-1-p.value.table
corrplot::corrplot(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], order="original", tl.pos="lt", method="color", tl.col="black",cl.lim=c(min(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]),max(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)])), is.corr=F,tl.cex=0.7, sig.level=(0.95),insig="pch", pch=5, p.mat=p_inv[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], pch.cex=0.15,pch.col="black",mar=c(3,1,5,1),cl.align.text="l")
pdf("corr_me9_vs_ce3_subsetprev_inv.pdf")
corrplot::corrplot(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], order="original", tl.pos="lt", method="color", tl.col="black",cl.lim=c(min(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)]),max(pb_mc.cor[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)])), is.corr=F,tl.cex=0.7, sig.level=(0.95),insig="pch", pch=5, p.mat=p_inv[1:ncol(era_gf), (ncol(era_gf)+1):ncol(pb_mc.cor)], pch.cex=0.15,pch.col="black",mar=c(3,1,5,1),cl.align.text="l")
dev.off()
```

```{r, fig.height=7, fig.width=15, echo=F}
pvclust.pb_mc.cor<-pvclust(pb_mc.cor, nboot = 1000)
pvclust.pb_mc.cor
plot(pvclust.pb_mc.cor)
pb_mc.cor.dist<-dist(pb_mc.cor)
plot(hclust(pb_mc.cor.dist))
```

