---
title: "Rueda-Alaña Chick Early Data"
author: "Rodrigo Senovilla Ganzo"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)
library(clustree)
library(pvclust)
library(Hmisc)
library(corrplot)
library(ggcorrplot)
library(stringr)
library(biomaRt)
```

## 1to1 otrhologs

Obtain orthologs names, the substitution step will be done at the end though

## Mouse genes 1to1

```{r cars}
if(exists("ensembl")==F){
ensembl<-useMart("ensembl")}
if(exists("mouse")==F){
mouse<-useDataset("mmusculus_gene_ensembl", mart=ensembl)}
attributes = c("external_gene_name", "ensembl_gene_id", "ggallus_homolog_ensembl_gene","ggallus_homolog_perc_id_r1", "ggallus_homolog_orthology_type", "ggallus_homolog_subtype", "ggallus_homolog_perc_id")
mouse_genes <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"), mart = mouse)
for(i in 1:length(mouse_genes[,1])){
  if(mouse_genes[i,1]==""){
    mouse_genes[i,1]<-mouse_genes[i,2]
  }
}
orth.mouse = getBM(attributes, filters="with_ggallus_homolog",values=TRUE, mart = mouse, bmHeader=FALSE)
orth.mouse<-orth.mouse[which(orth.mouse$ggallus_homolog_orthology_type=="ortholog_one2one"),]
ensembl.mouse<- orth.mouse[,2]
external.mouse<-orth.mouse[,1]
rownames(orth.mouse)<-ensembl.mouse
TFdata_mouse <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"),filters = 'go', values = c('GO:0006366', 'GO:0000981', 'GO:0003700', 'GO:0006383', 'GO:0000995', 'GO:0001228', 'GO:0001227'), mart = mouse)
```

## Chicken genes 1to1

```{r cars}
if(exists("ensembl")==F){
ensembl<-useMart("ensembl")}
datasets <- listDatasets(ensembl)
if(exists("chicken")==F){
chicken<-useDataset("ggallus_gene_ensembl", mart=ensembl)}
attributes = c("external_gene_name", "ensembl_gene_id", "mmusculus_homolog_ensembl_gene","mmusculus_homolog_perc_id_r1", "mmusculus_homolog_orthology_type", "mmusculus_homolog_subtype", "mmusculus_homolog_perc_id")
chicken_genes <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"), mart = chicken)
for(i in 1:length(chicken_genes[,1])){
  if(chicken_genes[i,1]==""){
    chicken_genes[i,1]<-chicken_genes[i,2]
  }
}
orth.chicken = getBM(attributes, filters="with_mmusculus_homolog",values=TRUE, mart = chicken, bmHeader=FALSE)
orth.chicken<-orth.chicken[which(orth.chicken$mmusculus_homolog_orthology_type=="ortholog_one2one"),]
ensembl.chicken<- orth.chicken[,2]
external.chicken<-orth.chicken[,1]
rownames(orth.chicken)<-ensembl.chicken
TFdata_chicken <- getBM(attributes=c('external_gene_name',"ensembl_gene_id"),filters = 'go', values = c('GO:0006366', 'GO:0000981', 'GO:0003700', 'GO:0006383', 'GO:0000995', 'GO:0001228', 'GO:0001227'), mart = chicken)
```



#Indentities

This step will be removed in the future, as the identities will be well stablished beforehand and we will know what each cluster stands for. 

```{r , echo=FALSE}
me9 <- readRDS("../mouse_e9/me9_sct5f.rds")
era <-readRDS("../chicken_e3/ce3_4f.rds")
```

```{r , echo=FALSE}
me9[['seurat_clusters']]<-me9[['me9_sct5_res2.8']]
era[['seurat_clusters']]<-era[['ce3_4']]
Idents(me9) <- "seurat_clusters"
Idents(era) <- "seurat_clusters"
```

```{r cars, warning=FALSE}
era_orth<-era
era.data<-GetAssayData(object = era, assay = "RNA")
era.data_ensembl<-merge(x=chicken_genes[,1:2], y=era.data, by.x= "external_gene_name", by.y= "row.names")
era.data_orth<-merge(x=orth.chicken[,2:3], y=era.data_ensembl, by.x= "ensembl_gene_id", by.y= "ensembl_gene_id")
era.data_orth<-era.data_orth[,-c(1,3)]
era.data_orth<-merge(x=mouse_genes, y=era.data_orth, by.x = "ensembl_gene_id", by.y = "mmusculus_homolog_ensembl_gene")
era.data_orth$external_gene_name<-sub('[.]','-',make.names(era.data_orth$external_gene_name, unique=T))
rownames(era.data_orth)<-toupper(era.data_orth$external_gene_name)
era.data_orth<-era.data_orth[,-c(1,2)]
era.data_orth<-Matrix(as.matrix(era.data_orth),sparse=TRUE)
mouse_gene_assay <- CreateAssayObject(era.data_orth)
era_orth[['RNA']] <- mouse_gene_assay
```


##Integration of datasets

```{r , echo=FALSE}
DefaultAssay(era_orth)<-"RNA"
chick.list <- SplitObject(era_orth, split.by="orig.ident")
chick.list <- lapply(X = chick.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("S.Score", "G2M.Score","percent.mt", "z.Score", "w.Score", "percent.hist"))
features <- SelectIntegrationFeatures(object.list = chick.list)
chick.list <- PrepSCTIntegration(object.list = chick.list, anchor.features = features)
chick.list <- lapply(X = chick.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = chick.list, reference = c(1, 2), reduction = "cca", dims = 1:50, normalization.method = "SCT", anchor.features = features)
era_orth <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```


```{r cars, warning=FALSE}
me9_orth<-me9
me9.data<-GetAssayData(object = me9, assay = "RNA")
orth.mouse[,1]<-toupper(orth.mouse[,1])
me9.data_orth<-merge(x=orth.mouse[,1:2], y=me9.data, by.x= "external_gene_name", by.y= "row.names")
rownames(me9.data_orth)<-me9.data_orth$external_gene_name
me9.data_orth<-me9.data_orth[,-c(1,2)]
me9.data_orth<-Matrix(as.matrix(me9.data_orth),sparse=TRUE)
mouse_gene_assay <- CreateAssayObject(me9.data_orth)
me9_orth[['RNA']] <- mouse_gene_assay
```

```{r}
DefaultAssay(me9_orth)<-"RNA"
me9_orth.list <- SplitObject(me9_orth, split.by = "orig.ident")
me9_orth.list <- lapply(X = me9_orth.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("percent.mt", "G2M.Score", "S.Score"))
features <- SelectIntegrationFeatures(object.list = me9_orth.list)
me9_orth.list <- PrepSCTIntegration(object.list = me9_orth.list, anchor.features = features)
me9_orth.list <- lapply(X = me9_orth.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = me9_orth.list, reduction = "cca", dims = 1:50, normalization.method = "SCT", anchor.features = features)
me9_orth <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```

# TRANSFERS

```{r}
me9_orth <- RunPCA(me9_orth, npcs = 50, verbose = FALSE, VariableFeatures(me9_orth),assay = "SCT")
me9_orth <- RunUMAP(me9_orth, dims = 1:50)
era_orth <- RunPCA(era_orth, npcs = 50, verbose = FALSE, VariableFeatures(era_orth),assay = "SCT")
era_orth <- RunUMAP(era_orth, dims = 1:50)
```



#Correlation matrix

Although the each species matrix only includes ortholog, we keep for comparisons only those genes presents in both datasets. 

#Chicken vs Mice

```{r, echo=F}
era_ct<- c(
"MB/DIEN-RP
HB-RP
HEM
AMP
OV
aTEL
bTEL
tHT
pHT
P3
aP2
bP2
P1
PG
aIT
bMB
pabMB
a'abMB
aMB
pIT
RL
aR1-2
aR3-8
abR0-8
bR0-8
DIEN-FP
MB-FP
IT-FP
HB-aFP
HB-bFP
TAL1 N
MOT N
TAL2 N
LHX9 N
LHX3 N
SPAL N
EI N
LI N"
)
era_ct<-str_split(era_ct, '\n')[[1]]
era_ct<-unlist(era_ct)
```

```{r, echo=F}
me9_ct<- c(
"MB/DIEN-RP
HB-RP
HEM
AMP
OV
TEL1
TEL2
SPAL
tHT
ATO
pHT
P3
P1-2
aIT
aMB
bMB
paIT
RL
EGL
aR1
aR2
aR3
aR4
aR5
abR1-2
abR3
abR4
bR0-2
bR3-5
MB-FP
IT-FP
HB-FP
MOT N
TAL2 N
MSX3 N
LHX N
LI N")
me9_ct<-str_split(me9_ct, '\n')[[1]]
me9_ct<-unlist(me9_ct)
```


# Mice in chicken

```{r , fig.height=10, fig.width=15, echo=FALSE}
Idents(me9_orth) <- "seurat_clusters"
Idents(era_orth) <- "seurat_clusters"
pb_anchors_minc <- FindTransferAnchors(query = me9_orth, reference = era_orth, dims = 1:50, reference.reduction = "pca", normalization.method = "SCT")
predictions <- TransferData(anchorset = pb_anchors_minc, refdata = era_orth$seurat_clusters, dims = 1:50)
me9_orth <- AddMetaData(me9_orth, metadata = predictions)
table(me9_orth$seurat_clusters,me9_orth$predicted.id)
era_orth <- RunUMAP(era_orth, dims = 1:50, reduction = "pca", return.model = TRUE)
me9_orth <- MapQuery(anchorset = pb_anchors_minc, reference = era_orth, query = me9_orth, refdata = list(celltype = "seurat_clusters"), reference.reduction = "pca", reduction.model = "umap")
p1_minc <- DimPlot(era_orth, reduction = "umap", group.by = "seurat_clusters", label = TRUE, label.size = 3, repel = TRUE) + ggtitle("Reference Mouse Annotations")+NoLegend()
me9_orth[["ref.umap"]]@cell.embeddings[,1]<-me9_orth[["ref.umap"]]@cell.embeddings[,1]*(-1)
p2_minc <- DimPlot(me9_orth, reduction = "ref.umap", group.by = "predicted.celltype", label = TRUE, label.size = 3, repel = TRUE) + ggtitle("Predicted Mouse Cell Types for Chicken Cells")+NoLegend()
p3_minc <- DimPlot(me9_orth, reduction = "ref.umap", group.by = "seurat_clusters", label = TRUE, label.size = 3, repel = TRUE) + ggtitle("Original Chicken Cell Types Transferred")+NoLegend()
```


```{r, }
p1_minc+p2_minc+p3_minc
pdf("mice_ref_minc.pdf")
p1_minc
dev.off()
pdf("predicted_chick_minc.pdf")
p2_minc
dev.off()
pdf("original_minc.pdf")
p3_minc
dev.off()
```


```{r, echo=F, fig.width=12, fig.height=12}
matches_minc<-as.matrix(table(me9_orth$seurat_clusters, me9_orth$predicted.id))
for(i in 1:nrow(matches_minc)){
  matches_minc[i,]<-matches_minc[i,]/sum(matches_minc[i,])
}
dim(matches_minc)
matches_minc<-matches_minc[me9_ct,]
lacking_c<-matrix(dimnames = list(me9_ct,era_ct[which(era_ct %nin% colnames(matches_minc))]), c(rep(0,length(me9_ct))))
matches_minc<-cbind(matches_minc, lacking_c)
matches_minc<-matches_minc[,era_ct]
dim(matches_minc)
pdf("matches_minc.pdf")
corrplot(matches_minc, method="shade", tl.col = "black", col.lim = c(0,1), col=colorRampPalette(c("blue","white","dark green"))(200))
corrplot(matches_minc, method="shade", tl.col = "black", col.lim = c(0,1))
corrplot(matches_minc, method="shade", tl.col = "black", col.lim = c(0,1), addgrid.col = "black")
dev.off()
?corrplot
#test 
## specialized the insignificant value according to the significant level
ggcorrplot(matches_minc, hc.order=T, lab =TRUE, insig = "blank")+ scale_fill_gradient2(limit =c(0,1), low = "white", high =  "black")
```

# Chicken in mice

```{r , fig.height=10, fig.width=15, echo=FALSE}
Idents(me9_orth) <- "seurat_clusters"
Idents(era_orth) <- "seurat_clusters"
pb_anchors_cinm <- FindTransferAnchors(query = era_orth, reference = me9_orth, dims = 1:50, reference.reduction = "pca", normalization.method = "SCT")
predictions <- TransferData(anchorset = pb_anchors_cinm, refdata = me9_orth$seurat_clusters, dims = 1:50)
era_orth <- AddMetaData(era_orth, metadata = predictions)
table(era_orth$seurat_clusters,era_orth$predicted.id)
me9_orth <- RunUMAP(me9_orth, dims = 1:50, reduction = "pca", return.model = TRUE)
era_orth <- MapQuery(anchorset = pb_anchors_cinm, reference = me9_orth, query = era_orth, refdata = list(celltype = "seurat_clusters"), reference.reduction = "pca", reduction.model = "umap")
p1_cinm <- DimPlot(me9_orth, reduction = "umap", group.by = "seurat_clusters", label = TRUE, label.size = 3, repel = TRUE) + ggtitle("Reference Chicken Annotations")+NoLegend()
p2_cinm <- DimPlot(era_orth, reduction = "ref.umap", group.by = "predicted.celltype", label = TRUE, label.size = 3, repel = TRUE) + ggtitle("Predicted Chicken Cell Types for Mouse Cells")+NoLegend()
p3_cinm <- DimPlot(era_orth, reduction = "ref.umap", group.by = "seurat_clusters", label = TRUE, label.size = 3, repel = TRUE) + ggtitle("Original Mouse Cell Types Transferred")+NoLegend()
```

```{r, echo=F}
p1_cinm+p2_cinm+p3_cinm
pdf("chicken_ref_cinm.pdf")
p1_cinm
dev.off()
pdf("predicted_mouse_cinm.pdf")
p2_cinm
dev.off()
pdf("original_cinm.pdf")
p3_cinm
dev.off()
```

```{r, echo=F, fig.width=12, fig.height=12}
matches_cinm<-as.matrix(table(era_orth$seurat_clusters,era_orth$predicted.id))
for(i in 1:nrow(matches_cinm)){
  matches_cinm[i,]<-matches_cinm[i,]/sum(matches_cinm[i,])
}
dim(matches_cinm)
matches_cinm<-matches_cinm[era_ct,]
n<-length(which(me9_ct %nin% colnames(matches_cinm)))
lacking_c<-matrix(rep(rep(0,length(era_ct)), n), ncol = n, dimnames = list(era_ct, me9_ct[which(me9_ct %nin% colnames(matches_cinm))]))
matches_cinm<-cbind(matches_cinm, lacking_c)
matches_cinm<-matches_cinm[,me9_ct]
dim(matches_cinm)
corrplot(matches_cinm, method="shade", tl.col = "black", col.lim = c(0,1), col=colorRampPalette(c("blue","white","dark green"))(200))
corrplot(matches_cinm, method="shade", tl.col = "black", col.lim = c(0,1), addgrid.col = "black")
```

```{r}
pdf("matches_cinm.pdf")
corrplot(matches_cinm, method="shade", tl.col = "black", col.lim = c(0,1), col=colorRampPalette(c("blue","white","dark green"))(200))
corrplot(matches_cinm, method="shade", tl.col = "black", col.lim = c(0,1), addgrid.col = "black")
dev.off()
```

#test 
## specialized the insignificant value according to the significant level
ggcorrplot(matches_cinm, hc.order=T, lab =TRUE, insig = "blank")+ scale_fill_gradient2(limit =c(0,1), low = "white", high =  "black")
```




















