---
title: "Rueda-Alaña Chick Early Data"
author: "Rodrigo Senovilla Ganzo"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)

```


```{r , echo=FALSE}
mouse_e9_integrate <- readRDS("me9_sct.rds")
me9_sct1 <- subset(mouse_e9_integrate, idents = c(5, 18, 21, 23, 26, 27), invert = TRUE)
```

```{r , echo=FALSE}
VlnPlot(me9_sct1, features = "nFeature_RNA")
VlnPlot(me9_sct1, features = "nCount_RNA")
VlnPlot(me9_sct1, features = "percent.mt")
plot1 <- FeatureScatter(me9_sct1, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(me9_sct1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```


To visualize the result, there are several ways to do it (VizDimReduction(), DimPlot(), and DimHeatmap())


```{r , echo=FALSE}
me9_sct1 <- RunPCA(me9_sct1, npcs = 50, verbose = FALSE)
print(me9_sct1[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(me9_sct1, dims = 1:2, reduction = "pca")

```

```{r , echo=FALSE}
DimPlot(me9_sct1, reduction = "pca")

```

```{r , echo=FALSE}
DimHeatmap(me9_sct1, dims = 1, cells = 500, balanced = TRUE)

```
This representation allows to explore primary sources of heterogenity in the dataset

```{r , echo=FALSE}
DimHeatmap(me9_sct1, dims = 15:30, cells = 500, balanced = TRUE)
for (i in 45:50) {
print(DimHeatmap(me9_sct1, dims = i, cells = 500, balanced = TRUE))}

```


## Alternative heuristic method for obtaining dimensionality

```{r , echo=FALSE}
ElbowPlot(me9_sct1, ndims = 50)

```

It is weird because it seems that all PC are important to the analysis regardin JackStraw, but considering Elbowplot, they are not!

## Cluster the cells 

Using different distance approach (euclidian distance in PCs)


```{r , echo=FALSE}
me9_sct1 <- FindNeighbors(me9_sct1, dims = 1:50)
me9_sct1 <- FindClusters(me9_sct1, resolution = 1)

```

An example for showing characterization of five initial cells

```{r , echo=FALSE}
head(Idents(me9_sct1), 5)

```

## Non-linear dimensional reduction (UMAP/tSNE)

This is the algorithm for UMAP

```{r , echo=FALSE}
me9_sct1 <- RunUMAP(me9_sct1, dims = 1:50)

```

and below its visualization



```{r , echo=FALSE}
DimPlot(me9_sct1, reduction = "umap")

```
This is the algorithm for tSNE

```{r , echo=FALSE}
me9_sct1 <- RunTSNE(me9_sct1, dims = 1:20)

```

and below its visualization


```{r , echo=FALSE}
DimPlot(me9_sct1, reduction = "tsne")

```

## Finding differentially expressed features (Cluster biomarkers) 

All cluster markers


```{r , echo=FALSE}
me9_sct1.markers <- FindAllMarkers(me9_sct1, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
for(i in unique(me9_sct1.markers$cluster)) {
  print(me9_sct1.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC) %>% filter(cluster == i))
}

top10<-me9_sct1.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
write.csv(top10 ,"me9_sct1_top10.csv")

```


## Visualization of gene expression across clusters



```{r , echo=FALSE}

VlnPlot(me9_sct1, features = c("HIST1H1E"))

```
```{r , echo=FALSE}

VlnPlot(me9_sct1, features = c("FGF8"))

```


```{r , echo=FALSE}
me9_sct1.gene.DE=(me9_sct1.markers %>% group_by(cluster) %>% top_n(n = 2, wt = avg_log2FC))$gene
for (i in 1:length(me9_sct1.gene.DE)){
print(FeaturePlot(me9_sct1, features = me9_sct1.gene.DE[i]))}

FeaturePlot(me9_sct1, features = "SLC17A6")
```

## 



```{r  echo=FALSE}
top10 <- me9_sct1.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
DoHeatmap(me9_sct1, features = top10$gene) + NoLegend()

```

## LEF1 



```{r , echo=FALSE}
new.cluster.ids <- c(0:(length(unique(me9_sct1$seurat_clusters))-1))
names(new.cluster.ids) <- levels(me9_sct1)
me9_sct1 <- RenameIdents(me9_sct1, new.cluster.ids)
p1<-DimPlot(me9_sct1, reduction = "umap", label = TRUE, pt.size = 0.5) 
p2<-DimPlot(me9_sct1, reduction = "tsne", label = TRUE, pt.size = 0.5) + NoLegend()
p1
c("MES","DTEL","HB","DMB","HB","DIEN","SUBPAL","WEIRD","ERY","ROOF","MES/FLOOR","ROM","ROM","SUBPAL","PRET","ITSHMUS","MB","DMB","DNTKNOW", "NCREST", "CHOROID", "MES/FLOOR","NOSENSE","PLATELETS")
```
```{r , echo=FALSE}
VlnPlot(me9_sct1, features = "nFeature_RNA")
VlnPlot(me9_sct1, features = "nCount_RNA")
VlnPlot(me9_sct1, features = "percent.mt")
plot1 <- FeatureScatter(me9_sct1, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(me9_sct1, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```

```{r , echo=FALSE}
saveRDS(me9_sct1, file = "./me9_sct1.rds")
```


