---
title: "Rueda-Alaña Chick Early Data"
author: "Rodrigo Senovilla Ganzo"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(Seurat)
library(SeuratObject)
library(patchwork)
library(Matrix)
library(glmGamPoi)
library(sctransform)
library(limma)
library(clustree)
library(pvclust)
library(readr)
library(readxl)
library(plyr)
```

```{r, echo=FALSE}
me9_sct5<-readRDS("me9_sct5i.rds")
```

```{r , echo=FALSE}
me9_sct5 <- RunPCA(me9_sct5, npcs = 50, verbose = FALSE)
print(me9_sct5[["pca"]], dims = 1:5, nfeatures = 5)
VizDimLoadings(me9_sct5, dims = 1:2, reduction = "pca")

```


```{r , echo=FALSE}
DimHeatmap(me9_sct5, dims = 1, cells = 500, balanced = TRUE)

```
This representation allows to explore primary sources of heterogenity in the dataset

```{r , echo=FALSE}
DimHeatmap(me9_sct5, dims = 15:30, cells = 500, balanced = TRUE)
for (i in 45:50) {
print(DimHeatmap(me9_sct5, dims = i, cells = 500, balanced = TRUE))}

```


## Alternative heuristic method for obtaining dimensionality

```{r , echo=FALSE}
ElbowPlot(me9_sct5, ndims = 50)

```

It is weird because it seems that all PC are important to the analysis regardin JackStraw, but considering Elbowplot, they are not!

## Cluster the cells 

Using different distance approach (euclidian distance in PCs)
An example for showing characterization of five initial cells

```{r , echo=FALSE}
resolution.range<-c(0.0,0.2,0.4,0.6,0.8,1,1.5,2,2.1,2.2,2.3,2.4,2.5,2.6,2.7,2.8,2.9,3,3.1,3.2,3.3,3.4)
me9_sct5 <- FindNeighbors(me9_sct5, dims = 1:50)
me9_sct5 <- Seurat::FindClusters(object = me9_sct5, resolution = resolution.range)
me9_sct5 <- RunUMAP(me9_sct5, dims = 1:50, resolution = resolution.range)
me9_sct5@meta.data
```
```{r , echo=FALSE}
Idents(me9_sct5)<-"integrated_snn_res.3.2"
DimPlot(me9_sct5, label = T)+NoLegend()
```


```{r , echo=FALSE}
Idents(me9_sct5)<-"me9_sct4_res2.5"
DimPlot(me9_sct5, label = T)+NoLegend()
```



Gene markers
## Finding differentially expressed features (Cluster biomarkers) 

All cluster markers


```{r, echo=FALSE}
curated_list<-read_xlsx("../Curated Gene Markers List.xlsx")
Idents(me9_sct5)<- "integrated_snn_res.3.2"
me9_sct5.markers <- FindAllMarkers(me9_sct5, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
me9_sct5.top10 <- me9_sct5.markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC)
me9_sct5.annot<- merge(me9_sct5.top10, curated_list, by.x= "gene") 
me9_sct5.annot<- rbind.fill(me9_sct5.annot, me9_sct5.top10[!(me9_sct5.top10$gene %in% me9_sct5.annot$gene),]) %>% arrange(cluster) %>% relocate(gene, .after= cluster)
me9_sct5.annot
```

## Visualization of gene expression across clusters



```{r, echo=F}
X<- me9_sct5.markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_log2FC) %>% filter(cluster==30)
for(i in 1:length(X$gene)){
  print(FeaturePlot(me9_sct5, X$gene[i]))
}
```

```{r, echo=F}
DefaultAssay(me9_sct5)<-"integrated"
Idents(me9_sct5)<-"integrated_snn_res.3.2"
me9_sct5.vs.markers <- FindMarkers(me9_sct5, ident.1 = 33, ident.2 = 0, only.pos = T, min.pct = 0.25, logfc.threshold = 0.25)
Y <- me9_sct5.vs.markers[1:70,]
for(i in 1:length(rownames(Y))) {
  print(FeaturePlot(me9_sct5, rownames(Y[i,])))
}
```

```{r, echo=F}
DefaultAssay(me9_sct5)<-"SCT"
H1<-rownames(me9_sct5[["SCT"]]@data)[grep("^FGF8", rownames(me9_sct5[["SCT"]]@data))]
for(i in 1:length(H1)){
  print(FeaturePlot(me9_sct5, features = H1[i]))}
```
```{r, echo=F}
MiceMarkers <- read.csv("../MiceMarkers.csv", sep = ";")
MiceMarkers <- c(MiceMarkers$X.)
MiceMarkers<-MiceMarkers %>% str_split(", ") %>% unlist()
DefaultAssay(me9_sct5)<-"SCT"
for(i in 1:length(MiceMarkers)){
  print(FeaturePlot(me9_sct5, features = MiceMarkers[i]))}
```

```{r, echo=F}
DefaultAssay(me9_sct5)<-"SCT"
FeaturePlot(me9_sct5, features = "DMRTA")
RidgePlot(me9_sct5, features = "RPL26", group.by = "Phase")
me9_sct5[["integrated"]]@var.features[grep("NKX", me9_sct5[["integrated"]]@var.features)]
```
## Subset neurons

```{r, warning=F}
Idents(me9_sct5)<- "integrated_snn_res.3.2"
me9_sct5.subset <- subset(me9_sct5, idents = c(19, 6, 25, 0, 33))
me9_sct5.subset <- RunPCA(me9_sct5.subset, features = VariableFeatures(object = me9_sct5.subset))
me9_sct5.subset <- FindNeighbors(me9_sct5.subset, dims = 1:50)
me9_sct5.subset <- FindClusters(me9_sct5.subset, resolution = resolution.range)
me9_sct5.subset <- RunUMAP(me9_sct5.subset, dims = 1:50, resolution = resolution.range)
```
```{r}
Idents(me9_sct5.subset)<-"integrated_snn_res.0.4"
FeaturePlot(me9_sct5.subset, features = c("TUBB6","COL1A1", "FABP7"))
DimPlot(me9_sct5.subset, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
Idents(me9_sct5.subset)<-"me9_sct5_res2.8"
DimPlot(me9_sct5.subset, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

```{r, echo=F}
Idents(me9_sct5.subset)<-"integrated_snn_res.0.4"
me9_sct5.subset.markers <- FindAllMarkers(me9_sct5.subset, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 2)
me9_sct5.subset.top10 <- me9_sct5.subset.markers %>% group_by(cluster) %>% top_n(n = 30, wt = avg_log2FC)
me9_sct5.subset.annot<- merge(me9_sct5.subset.top10, curated_list, by.x= "gene") 
me9_sct5.subset.annot<- rbind.fill(me9_sct5.subset.annot, me9_sct5.subset.top10[!(me9_sct5.subset.top10$gene %in% me9_sct5.subset.annot$gene),]) %>% arrange(cluster) %>% relocate(gene, .after= cluster)
me9_sct5.subset.annot
```


```{r, echo=F}
for(i in 1:length(me9_sct5.subset.top10$gene)) {
  print(FeaturePlot(me9_sct5.subset, me9_sct5.subset.top10$gene[[i]]))
}
```

```{r, echo=F}
for(i in 1:length(me9_sct5.subset.top10$gene)) {
  print(FeaturePlot(me9_sct5, me9_sct5.subset.top10$gene[[i]]))
}
```
```{r,  echo=FALSE}
top10 <- me9_sct5.markers %>% group_by(cluster) %>% top_n(n = 10, wt = avg_log2FC)
DoHeatmap(me9_sct5, features = top10$gene) + NoLegend() + theme(text = element_text(size = 20))
```

## Cluster annnotation

```{r , echo=FALSE}
me9_sct5<-readRDS("./me9_sct5f.rds")

Idents(me9_sct5)<-"integrated_snn_res.2.8"
DimPlot(me9_sct5, label = T)+NoLegend()
new.cluster.ids <-c("aMB", "TEL1", "TEL2", "OV", "P1-2", "abR1-2", "abR3", "aR1", "HEM",  "aR4", "aIT", "aR3", "bR3-5", "SPAL", "paIT", "LHX N", "HB-FP", "bMB", "TAL2 N", "HB-RP", "P3", "bR0-2", "aR5", "IT-FP", "MB-FP", "tHT", "RL", "MB/DIEN-RP", "abR4", "pHT", "MOT N", "LI N", "MSX3 N", "ATO", "AMP", "EGL", "aR2")
names(new.cluster.ids) <- levels(me9_sct5)
me9_sct5 <- RenameIdents(me9_sct5, new.cluster.ids)
me9_sct5[["me9_sct5_res2.8"]]<-Idents(me9_sct5)
p1<-DimPlot(me9_sct5, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
p1
```
```{r}
pdf("UMAP_me9_sct5.pdf")
p1
dev.off()
```

```{r , echo=FALSE}
saveRDS(me9_sct5, file = "./me9_sct5f.rds")
```

##SUBSETTING



{r , echo=FALSE}
Idents(me9_sct5)<-"integrated_snn_res.3"
DimPlot(me9_sct5, label=T)
me9_sct5 <- subset(me9_sct5, idents = c(35), invert = TRUE)
```

```{r , echo=FALSE}
VlnPlot(me9_sct5, features = "nFeature_RNA")
VlnPlot(me9_sct5, features = "nCount_RNA")
VlnPlot(me9_sct5, features = "percent.mt")
plot1 <- FeatureScatter(me9_sct5, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(me9_sct5, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```
```{r, echo=F}
library(readr)
cc.fer <- read_delim("../Cellcycle/fer_genes.csv", 
    ";", escape_double = FALSE, trim_ws = TRUE)
s.genes <- unique(toupper(c(cc.genes$s.genes,cc.fer$`Phase S`)))
g2m.genes <- unique(toupper(c(cc.genes$g2m.genes,cc.fer$`Phase G2M`)))
```


To visualize the result, there are several ways to do it (VizDimReduction(), DimPlot(), and DimHeatmap())

```{r}
DefaultAssay(me9_sct5)<-"RNA"
me9_sct5.list <- SplitObject(me9_sct5, split.by = "orig.ident")
me9_sct5.list <- lapply(X = me9_sct5.list, FUN = SCTransform, method = "glmGamPoi",  vars.to.regress = c("percent.mt", "G2M.Score", "S.Score"))
features <- SelectIntegrationFeatures(object.list = me9_sct5.list)
me9_sct5.list <- PrepSCTIntegration(object.list = me9_sct5.list, anchor.features = features)
me9_sct5.list <- lapply(X = me9_sct5.list, FUN = RunPCA, features = features)
anchors <- FindIntegrationAnchors(object.list = me9_sct5.list, reduction = "cca", dims = 1:50, normalization.method = "SCT", anchor.features = features)
me9_sct5 <- IntegrateData(anchorset = anchors, normalization.method = "SCT", dims = 1:50)
```

```{r , echo=FALSE}
saveRDS(me9_sct5, file = "./me9_sct6i.rds")
```
